### 2024-09-23 outputs/paper/paper.qmd
created 4 figures for the data section. llm usage updated

---
title: "My title"
# Descriptive, informative, specific title
subtitle: "My subtitle if needed"
author: 
  - Lexi Knight
thanks: "Code and data are available at: https://github.com/LexiKnight/neighborhoods_toronto/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
# Tell reader top level findings
# One thing we learn about the world because of this paper?
# Abstracts need to be “tightly written”, almost terse. Remove unnecessary words. Do not include more than four sentences.

format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

I use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

Notes
2-3 paragraphs of content
Additional paragraph that sets out the remainder of the paper. telegraph the rest of the paper: “Section 2…, Section 3….”.
In the introduction, please be more specific about your findings
Example
The remainder of this paper is structured as follows. First, the methods used to analyze global and local patterns of segregation are reviewed. Next, the data and study region are described, the multilevel models are detailed, and the results are presented. Finally, the contributions of this study for understanding residential income segregation are discussed, the limitations are highlighted, and directions for future research are proposed

# Data {#sec-data}

```{r}

# install libraries 
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("ggplot2")
# install.packages("here")

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)

# Check if the file exists and load it 
file_path <- here::here("inputs/data/analysis_data/neighborhood_analysis_data.csv")

if (file.exists(file_path)) {
  analysis_data <- read_csv(file_path)
  print("File found and loaded successfully!")
} else {
  stop("Error: File not found at the specified path.")
}

# Reshape the data to long format for easier plotting
long_data <- analysis_data %>%
  pivot_longer(cols = -`Neighbourhood Name`, 
               names_to = "Neighborhood", 
               values_to = "Value")

## Graphing visible minority data
# Define categories
visible_minority_categories <- c("South Asian", "Chinese", "Black", "Filipino",
                                 "Arab", "Latin American", "Southeast Asian",
                                 "West Asian", "Korean", "Japanese",
                                 "Visible minority, n.i.e.",
                                 "Multiple visible minorities",
                                 "Not a visible minority")

# Filter for visible minority categories
filtered_data <- long_data %>%
  filter(`Neighbourhood Name` %in% visible_minority_categories)

# Calculate total individuals for each neighborhood
total_individuals <- filtered_data %>%
  group_by(Neighborhood) %>%
  summarize(Total = sum(Value), .groups = "drop")  # Use .groups to avoid grouping issues

# Create the plot with number of individuals
visible_minorities_plot <- ggplot(filtered_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Visible Minority Categories Between Neighborhoods",
       x = "Visible Minority Categories",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "lightgreen", "Downtown Yonge East" = "forestgreen")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_data$Value), by = 2000))  # Adjust 'by' value as needed

# Display the plot
print(visible_minorities_plot)

## Graphing income data
# Define categories
income_categories <- c("Under $10,000", "$10,000 to $19,999",
                       "$20,000 to $29,999", "$30,000 to $39,999",
                       "$40,000 to $49,999", "$50,000 to $59,999",
                       "$60,000 to $69,999", "$70,000 to $79,999",
                       "$80,000 to $89,999", "$90,000 to $99,999",
                       "$100,000 to $149,999", "$150,000 and over")

# Convert the 'Neighbourhood Name' to a factor with specified levels for ordering
filtered_income_data$`Neighbourhood Name` <- factor(filtered_income_data$`Neighbourhood Name`, levels = income_categories)

# Create the plot with number of individuals for income
income_plot <- ggplot(filtered_income_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Income Categories Between Neighborhoods",
       x = "Income Categories",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "magenta", "Downtown Yonge East" = "darkmagenta")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_income_data$Value), by = 500))  # Adjust 'by' value for more tick marks

# Display the plot
print(income_plot)

## Graphing education data
# Define the ordered education categories
education_categories <- c("No certificate, diploma or degree",
                          "High school diploma or equivalency certificate", 
                          "Postsecondary certificate or diploma",
                          "Apprenticeship or trades certificate or diploma",
                          "Non-apprenticeship trades certificate or diploma",
                          "College, CEGEP or other non-university certificate or diploma",
                          "University certificate or diploma below bachelor level",
                          "Bachelor's degree",
                          "University certificate or diploma above bachelor level",
                          "Degree in medicine, dentistry, veterinary medicine or optometry",
                          "Master's degree", "Earned doctorate")

# Filter the long_data for education categories
filtered_education_data <- long_data %>%
  filter(`Neighbourhood Name` %in% education_categories)

# Convert 'Neighbourhood Name' to a factor for ordered plotting
filtered_education_data$`Neighbourhood Name` <- factor(filtered_education_data$`Neighbourhood Name`, levels = education_categories)

# Create the plot with number of individuals for education attainment
education_plot <- ggplot(filtered_education_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Education Attainment Between Neighborhoods",
       x = "Education Levels",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "skyblue", "Downtown Yonge East" = "blue")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_education_data$Value), by = 1000))  # Adjust 'by' value for more tick marks

# Display the plot
print(education_plot)

## Graphing knowledge of official language
# Define the categories for language knowledge
language_categories <- c("English only", "French only", 
                         "English and French", "Neither English nor French")

# Filter the long_data for language knowledge categories
filtered_language_data <- long_data %>%
  filter(`Neighbourhood Name` %in% language_categories)

# Convert 'Neighbourhood Name' to a factor for ordered plotting
filtered_language_data$`Neighbourhood Name` <- factor(filtered_language_data$`Neighbourhood Name`, levels = language_categories)

# Create the plot with number of individuals for knowledge of official languages
language_plot <- ggplot(filtered_language_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Knowledge of Official Languages Between Neighborhoods",
       x = "Knowledge of Official Languages",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "yellow", "Downtown Yonge East" = "gold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_language_data$Value), by = 2000))  # Adjust 'by' value for more tick marks

# Display the plot
print(language_plot)

```

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes).

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

\newpage

# References